---
title: "Titanic"
output:
  html_document: default
  pdf_document: default
date: "2024-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The data-set

Data Dictionary

Variable	
survival:	Survival	0 = No, 1 = Yes

pclass:	Ticket class	1 = 1st, 2 = 2nd, 3 = 3rd

sex:	Sex	

Age:	Age in years	

sibsp:	# of siblings / spouses aboard the Titanic

parch:	# of parents / children aboard the Titanic

ticket:	Ticket number	

fare:	Passenger fare

cabin	:Cabin number	

embarked:	Port of Embarkation	C = Cherbourg, Q = Queenstown, S = Southampton

pclass: A proxy for socio-economic status (SES)

1st = Upper
2nd = Middle
3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...

Sibling = brother, sister, stepbrother, stepsister

Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...

Parent = mother, father

Child = daughter, son, stepdaughter, stepson

Some children travelled only with a nanny, therefore parch=0 for them.

# TRAIN DATA
```{r,comment=NA,echo=FALSE}
library("rmarkdown")
data=read.csv("C:/Users/HP/Downloads/train.csv")
paged_table(data)
```

# Dimension of the data-set

```{r,comment=NA,echo=FALSE}
dim(data)
```

# Variables in the data-set

```{r,comment=NA,echo=FALSE}
str(data)
```

# Column Names

```{r,comment=NA,echo=FALSE}
colnames(data)
```

### **We remove the 1st,4th,9th and 11th column from the data-set**

```{r,comment=NA,echo=FALSE}
ndata=data[-c(1,4,9,11)]
```

### **We replace the NA values in the Age column with the mean of that column and convert the character variables into factor variables.**

```{r,comment=NA,echo=FALSE}
age.mean=mean(ndata$Age,na.rm=TRUE)
ndata$Age=replace(ndata$Age,is.na(ndata$Age)==1,age.mean)
ndata$Sex=as.factor(data$Sex)
ndata$Embarked=as.factor(data$Embarked)
ndata$Pclass=as.factor(data$Pclass)
```

### **In the dataset, we remove the empty cells from the 'Embark' column**

```{r,comment=NA,echo=FALSE}
index=which(ndata$Embarked=="")
ndata=ndata[-index,]
```

# PLOTS
```{r,echo=FALSE}
library("ggplot2")
#Survived
data=data.frame("Category"=c("0","1"),"values"=c(sum(ndata$Survived=="0"),sum(ndata$Survived=="1")))
ggplot(data,aes(x=Category,y=values,fill=Category))+geom_bar(stat="identity")+geom_text(aes(label=values))+scale_fill_manual(values=c("orange","blue"))+scale_x_discrete(limits =c("0","1"))+labs(title="Survived")
#Pclass
data=data.frame("Category"=c("1","2","3"),"values"=c(sum(ndata$Pclass=="1"),sum(ndata$Pclass=="2"),sum(ndata$Pclass=="3")))
ggplot(data,aes(x=Category,y=values,fill=Category))+geom_bar(stat="identity")+geom_text(aes(label=values))+scale_fill_manual(values=c("green","#993300","#666666"))+scale_x_discrete(limits =c("1","2","3"))+labs(title="Pclass")
#Sex
data1=data.frame("cat1"=c("female","male"),"val1"=c(sum(ndata$Sex=="female"),sum(ndata$Sex=="male")))
slices1=c(sum(ndata$Sex=="female"),sum(ndata$Sex=="male"))
frac1=(slices1/sum(slices1))
ymax1=cumsum(frac1)
ymin1=c(0,head(ymax1,n=-1))
labposi1=(ymax1+ymin1)/2
labls1=paste0(c("female","male"),"\n value:",paste(round(frac1*100)),"%",sep="")
ggplot(data1,aes(ymax=ymax1,ymin=ymin1,xmax=4,xmin=3,fill=cat1))+geom_rect()+geom_label(x=3.5,aes(y=labposi1,label=labls1),size=3)+coord_polar(theta="y")+xlim(c(2,4))+theme_void()+theme(legend.position = "none")+labs(title=" Sex")+scale_fill_manual(values=c("yellow","purple"))
#Age

ggplot(data=ndata,aes(x=Age))+geom_histogram(aes(y=..density..),bins=24,col="black",fill="#00FF99")+labs(title="Age")+geom_density()
#Fare
ggplot(data=ndata,aes(x=Fare))+geom_histogram(aes(y=..density..),bins=24,col="black",fill="#66CCFF")+labs(title="Fare")+geom_density()
#Parch
ggplot(data=ndata,aes(x=Parch))+geom_histogram(aes(y=..density..),bins=24,col="black",fill="#FFFF99")+labs(title="Parch")+geom_density()



#Embarked
data=data.frame("Category"=c("C","Q","S"),"values"=c(sum(ndata$Embarked=="C"),sum(ndata$Embarked=="Q"),sum(ndata$Embarked=="S")))
ggplot(data,aes(x=Category,y=values,fill=Category))+geom_bar(stat="identity")+geom_text(aes(label=values))+scale_fill_manual(values=c("green","#993300","yellow"))+scale_x_discrete(limits =c("C","Q","S"))+labs(title=" Embarked",x="",y="Count")



```

### ***Since our response variable is binary, taking only the values 0 and 1, we use a logistic regression model.***

```{r,comment=NA,echo=FALSE}
model=glm(Survived~Pclass+Sex+Age+SibSp+Parch+Fare+Embarked,family="binomial",data=ndata)
summary(model)
```

*We can see that 'Pclass2', 'Pclass3', 'SexMale', 'Age', and 'SibSp' are statistically significant, as their p-values are less than 0.05.*

*To improve the fit of our model we use AIC backward method.* 

### ***Backward elimination based on the Akaike Information Criterion(AIC)***

```{r,comment=NA,,echo=FALSE}
model2=step(model,direction="backward")
```

*We have selected the variables 'Embarked', 'SibSp', 'Age', 'Pclass', and 'Sex' for our regression model.*

### **Half-Normal Probability (hnp) Plot**

```{r,comment=NA,echo=FALSE}
library("hnp")
hnp(model2)
```

*Most of the residuals fall within these confidence bands, suggesting that the model is capturing the majority of the variability in the data accurately.*

## **Residual vs Fitted Plot**

```{r,comment=NA,echo=FALSE}
plot(residuals(model2,"pearson"),main="Residuals Plot",xlab="Fitted Values",ylab="Pearson Residuals")
```

### **Receiver Operator Characteristic (ROC) Curve and Area Under the Curve (AUC)**

```{r,comment=NA,echo=FALSE}
library("caret")
library("ROCR")
pred=predict(model2,type="response",newdata=ndata)
pred1=prediction(pred,ndata$Survived)
pref=performance(pred1,"tpr","fpr")
plot(pref,print.cutoffs.at=seq(0,1,0.1),colorize=TRUE) 
abline(a=0,b=1)
auc=performance(pred1,"auc")@y.values[[1]]
legend(0.6,0.4,auc,title="AUC",cex=0.9)
```

*The model's AUC is 0.8567, indicating good predictive performance.*

# TEST DATA

```{r,comment=NA,echo=FALSE}
dat=read.csv("C:\\Users\\HP\\Downloads\\test.csv")
paged_table(dat)
```


```{r,comment=NA,echo=FALSE}
ndat=dat[-c(1,4,9,11)]
age.mean=mean(ndat$Age,na.rm=TRUE)
ndat$Age=replace(ndat$Age,is.na(ndat$Age)==1,age.mean)
ndat$Sex=as.factor(dat$Sex)
ndat$Embarked=as.factor(dat$Embarked)
ndat$Pclass=as.factor(dat$Pclass)

```


# PREDICTION

```{r,comment=NA,echo=FALSE}
predi=predict(model2,ndat)
val=ifelse((predi)>0.5,1,0)
r=data.frame("PassengerId"=dat[,1],"Survived"=val)
paged_table(r)
```

